---
title: 'PROJECT REPORT'
subtitle: 'FINANCIAL TRADING STRATEGY'
author: 'Diane Kamning'
date: March 22, 2017
output:
  pdf_document:
    toc: true
    number_sections: true
mainfont: Times New Roman
fontsize: 12pt
header-includes:
  - \setlength\parindent{24pt}
---

\pagebreak


```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE }
#### Load libraries ####
library(rmarkdown)
library(zoo)
library(xts)
library(TTR)
library(quantmod)

library(dplyr)
#library(data.table)

library(FinancialInstrument)
library(PerformanceAnalytics)
library(blotter)
library(Quandl)
library(foreach)
library(quantstrat)

## Packages for nice table display
pkg <- c("dplyr", "knitr", "devtools", "DT", "xtable")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
    install.packages(new.pkg, repos = "http://cran.rstudio.com")
}
# Load the packages into R
library(dplyr)
library(knitr)
library(DT)
library(xtable)


#### Load scripts ####
#getwd()
#setwd('C:/Users/d_kamning/OneDrive for Business/Private/Springboard/Capstone')
source('C:/Users/d_kamning/OneDrive for Business/Private/Springboard/Capstone/Final/functions_util.R')


#### LOAD DATA ####
### stockOne is AEP while stockTwo is CHK (Chesapeake Energy Corp.) ###
stockOne <- Quandl("WIKI/AEP", start_date = "2000-01-01", end_date = "2017-01-31", type = "xts")
head(stockOne)
stockOne <- stockOne[, c(1, 2, 3, 4, 5, 6)]


stockTwo <- Quandl("WIKI/CHK", start_date = "2000-01-01", end_date = "2017-01-31", type = "xts")
head(stockTwo)
stockTwo <- stockTwo[, c(1, 2, 3, 4, 5, 6)]
```

# INTRODUCTION  
  
  This project aims at building a model that will ideally always output successful bids in the stock market. For that, it builds a model which gives better results when constantly trained in a sliding-time window.
The goal is to design a simple financial trading strategy that will be profitable and that will provide a good risk-adjusted measure of return.  


# DATA SETS  
  
  Two datasets will be used here to test the strategy:

- The American Electric Company (AEP) dataset from Quandl
```{r, echo=FALSE, message = FALSE}
knitr::kable(head(stockOne) )
```  
  
- The Chesapeake Energy Corporation (CHK) from Quandl.
```{r, echo=FALSE, message = FALSE}
knitr::kable(head(stockTwo))
```
  
  An initial exploration of the AEP dataset reveals 5 important fields:  

- The date
- The Open price
- The High price
- The Low price
- The Close price.  
  
  Some of the issues encountered with the data:  

- The presence of the adjusted closing price was confusing for some methods in the packages Quandstrat and xts, as those methods kept throwing errors. I had to remove the adjusted closing price from my data sets and just keep the closing price
- Some functions and arguments were not found because the Quandstrat package is not yet stable.


# PRELIMINARY EXPLORATION  
  
  Indicators are transformations of market data that give an insight into the overall market behavior by measuring current conditions and/or forecasting trends.  Among others, there are trend-following indicators which depict the general price direction, and oscillators used to discover on a scale of 0 to 100 short-term overbought (above 70 to 80) or oversold (below 30 to 20) conditions .
  Combining trend-following indicators and oscillator/reversion indicators gives more insight into the data for this project. The preliminary oscillator used is an RSI (Relative Strength Index) with a 3-days lookback period. The preliminary trend indicators are 3 SMA (Simple Moving Average). After applying those indicators to the stocks, there are some periods of time during which none of the indicators seem to be right.  


## TREND-FOLLOWING INDICATORS: SIMPLE MOVING AVERAGES  
  
  The SMA50 (Simple Moving Average) seems to better mimic the trend of the closing prices for both data sets

- SMA AEP

```{r, echo=FALSE}
#### VISUALIZE DATA ####
### Stock 1 SMA
plot(Cl(stockOne), type = 'l', xlab = 'Date', ylab = 'Closing price', main = 'AEP closing price trend with SMA')
lines(SMA(Cl(stockOne), n = 200), col = "red")
lines(SMA(Cl(stockOne), n = 300), col = "purple")
lines(SMA(Cl(stockOne), n = 50), col = "cyan")
legend('topleft', c("SMA200", "SMA300", "SMA50", "Closing price"),
    lty = c(1, 1), lwd = c(2.5, 2.5), col = c('red', 'purple', 'cyan', "black"), cex = .50)
```
  
- SMA CHK

```{r, echo=FALSE}
#### VISUALIZE DATA ####
### Stock 2 SMA
plot(Cl(stockTwo), type = 'l', xlab = 'Date', ylab = 'Closing price', main = 'CHK closing price trend with SMA')
lines(SMA(Cl(stockTwo), n = 200), col = "red")
lines(SMA(Cl(stockTwo), n = 300), col = "purple")
lines(SMA(Cl(stockTwo), n = 50), col = "cyan")
legend('topleft', c("SMA200", "SMA300", "SMA50", "Closing price"),
    lty = c(1, 1), lwd = c(2.5, 2.5), col = c('red', 'purple', 'cyan', "black"), cex = .50)
```



## OSCILLATOR/REVERSION INDICATOR: RSI
    
  An observation of the graphs of the stocks’ RSI reveals that there are effectively periods of reversion (2013-09-03 to 2013-9-05 for example) that won’t be captured by a trend-following indicator:

- RSI AEP

```{r, echo=FALSE}
#### VISUALIZE DATA ####
### Stock 1 RSI
plot(RSI(Cl(stockOne), n = 3), type = "l", xlab = 'Date', ylab = 'Closing price', main = 'AEP closing price RSI with 3-days lookback')
```

- RSI CHK

```{r, echo=FALSE}
#### VISUALIZE DATA ####
### Stock 2 RSI
plot(RSI(Cl(stockTwo), n = 3), type = "l", xlab = 'Date', ylab = 'Closing price', main = 'CHK closing price RSI with 3-days lookback')
```


# APPROACH
  
  The main objective is to obtain a profit factor above 1 after running the strategy on each of the data sets. 
The approach here would be to combine both SMA50 and SMA200 with an oscillator to avoid false signals. The trend-following indicators would help catch up trends.  
	The analysis is done on 13 years, from January 01s, 2013 to December 31st, 2016 . The time is Eastern time and the currency is US dollar.

```{r, echo=FALSE, message=FALSE, include=FALSE}
#### INITIALIZE STRATEGY ####
### Create initdate, from, and to charater strings
initdate <- "2000-01-01"
from <- "2003-01-01"
to <- "2016-12-31"

### Set the timezone to UTC
#OlsonNames()
Sys.setenv(TZ = "EST")

### Set the currency to USD 
currency("USD")

### Use the stock command to initialize and set currency to USD
stock(c("stockOne", "stockTwo"), currency = "USD")


### Define your trade size and initial equity
tradesize <- 100000
initeq <- 100000

### Define the names of your strategy, portfolio and account
strategyRSI.st <- "RSI strategy"
portfolioRSI.st <- "RSI strategy"
accountRSI.st <- "RSI strategy"

### Remove the existing strategy if it exists
rm.strat(strategyRSI.st)


### initialize the portfolio
initPortf(portfolioRSI.st, symbols = c("stockOne", "stockTwo"), initDate = initdate, currency = "USD")

### initialize the account
initAcct(accountRSI.st, portfolios = portfolioRSI.st, initDate = initdate, currency = "USD", initEq = initeq)

### initialize the orders
initOrders(portfolioRSI.st, initDate = initdate)

### store the strategy
strategy(strategyRSI.st, store = TRUE)
```  
  

## STRATEGY 1: RSI  

  "RSI Strategy", the initial strategy for this project, uses simple averages (over 50 days and over 200 days) with a custom RSI_3_4 indicator acting as an average between RSI3 and RSI4 . 
  Signals help interpret how indicators interact with the market and with each other. Those signals are:  

- a comparison and a crossover , which show a buy signal when the 50-day simple moving average is above the 200-day simple moving average and show a sell signal when the 50-day simple moving average crosses below the 200-day simple moving average
- a threshold, which an oversold condition, thus a buy opportunity, for RSI_3_4 below 20, and an overbought condition, thus a sell opportunity, for RSI_3_4 above 80
- a combined comparison and threshold to buy when the 50-day simple moving average is above the 200-day simple moving average and RSI_3_4 is less then 20.  

```{r, echo=FALSE, results='hide'}
#### ADD SMA INDICATORS (TREND) ####
### 1. Add a 200-day simple moving average indicator to your strategy
add.indicator(strategy = strategyRSI.st,
# Add the SMA function
                  name = "SMA",
# Create a lookback period
                  arguments = list(x = quote(Cl(mktdata)), n = 200),
# Label your indicator SMA200
                  label = "SMA200")


### 2. Add a 50-day simple moving average indicator to your strategy
add.indicator(strategy = strategyRSI.st,
# Add the SMA function
                  name = "SMA",
# Create a lookback period
                  arguments = list(x = quote(Cl(mktdata)), n = 50),
# Label your indicator SMA50
                  label = "SMA50")



##### ADD CUSTOM RSI INDICATOR (REVERSION) ####
### Add the RSI_avg function to your strategy using an n1 of 3 and an n2 of 4, and label it "RSI_3_4"
add.indicator(strategyRSI.st, name = "RSI_avg", arguments = list(price = quote(Cl(mktdata)), n1 = 3, n2 = 4), label = "RSI_3_4")


#### ADD OSCILLATOR DVO (David Varadi Oscillator, originated by David Varadi) ####
### Add the DVO indicator to your strategy
#add.indicator(strategy = strategyRSI.st, name = "DVO",
              #arguments = list(HLC = quote(HLC(mktdata)), navg = 2, percentlookback = 126),
              #label = "DVO_2_126")


#### ADD SIGNALS ####
### 1. Using sigComparison ################
# add a sigComparison which specifies that SMA50 must be greater than SMA200, call it longfilter
add.signal(strategyRSI.st, name = "sigComparison",
# we are interested in the relationship between the SMA50 and the SMA200
               arguments = list(columns = c("SMA50", "SMA200"),
# particularly, we are interested when the SMA50 is greater than the SMA200
                                relationship = "gt"),
# label this signal longfilter
               label = "longfilter")

### 2. Using sigCrossover ###########
# add a sigCrossover which specifies that the SMA50 is less than the SMA200 and label it filterexit
add.signal(strategyRSI.st, name = "sigCrossover",
# we're interested in the relationship between the SMA50 and the SMA200
               arguments = list(columns = c("SMA50", "SMA200"),
# the relationship is that the SMA50 crosses under the SMA200
                                relationship = "lt"),
# label it filterexit
               label = "filterexit")


### 3.a Using sigThreshold - I ########
# SigThreshold() function takes the cross argument, which specifies whether it will function similarly to sigComparison (cross = FALSE) or sigCrossover (cross = TRUE), respectively.
# Implement a sigThreshold which specifies that RSI_3_4 must be less than 20, label it longthreshold
add.signal(strategyRSI.st, name = "sigThreshold",
# use the RSI_3_4 column
               arguments = list(column = "RSI_3_4",
# the threshold is 20
                                threshold = 20,
# we want the oscillator to be under this value
                                relationship = "lt",
# we're interested in every instance that the oscillator is less than 20
                                cross = FALSE),
# label it longthreshold
               label = "longthreshold")


### 3.b Using sigThreshold - II ###########
# add a sigThreshold signal to your strategy that specifies that RSI_3_4 must cross above 80 and label it thresholdexit
add.signal(strategyRSI.st, name = "sigThreshold",
# reference the column of RSI_3_4
               arguments = list(column = "RSI_3_4",
# set a threshold of 80
                                threshold = 80,
# the oscillator must be greater than 80
                                relationship = "gt",
# we are interested only in the cross
                                cross = TRUE),
# label it thresholdexit
               label = "thresholdexit")



### 4. Using sigFormula, Combining signals ############
# add a sigFormula signal to your code specifying that both longfilter and longthreshold must be TRUE, label it longentry
add.signal(strategyRSI.st, name = "sigFormula",
# specify that longfilter and longthreshold must be TRUE
               arguments = list(formula = "longfilter & longthreshold",
# specify that cross must be TRUE
                                cross = TRUE),
# label it longentry
               label = "longentry")

# Test and Subset your data between Sep. 1 and Sep. 5 of 2013
test_init <- applyIndicators(strategyRSI.st, mktdata = OHLC(stockOne))
test <- applySignals(strategy = strategyRSI.st, mktdata = test_init)
knitr::kable(head(test))
test_subset <- test["2013-09-01/2013-09-05"]
knitr::kable(head(test_subset))
 
test_init2 <- applyIndicators(strategyRSI.st, mktdata = OHLC(stockTwo))
test2 <- applySignals(strategy = strategyRSI.st, mktdata = test_init2)
knitr::kable(head(test2))
test_subset2 <- test2["2013-09-01/2013-09-05"]
knitr::kable(head(test_subset2))
```
  
  Rules help shape trading transactions at signal execution. They generate orders using market data, indicators and signals. This strategy has 2 rules:  

- an entry rule of 1 share for the combined comparison and threshold entry signals
- an exit rule for a treshold above 80.

```{r, echo=FALSE, results = 'hide'}
#### ADD RULES ####
### Using add.rule() to implement an exit rule #########
add.rule(strategyRSI.st, name = "ruleSignal",
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all",
                          ordertype = "market", orderside = "long",
                          replace = FALSE, prefer = "Open"),
         type = "exit")

#In quantstrat, the replace argument specifies whether or not to ignore all other signals on the same date when the strategy acts upon one signal.
### In this case, you will be working with a rule that sells when the oscillator has crossed a certain threshold. In particular, you will be working with the thresholdexit rule now.
add.rule(strategyRSI.st, name = "ruleSignal",
         arguments = list(sigcol = "thresholdexit", sigval = TRUE, orderqty = "all",
                          ordertype = "market", orderside = "long",
                          replace = FALSE, prefer = "Open"),
         type = "exit")


### Using add.rule() to implement an entry rule #########
# create an entry rule of 1 share when all conditions line up to enter into a position
add.rule(strategyRSI.st, name = "ruleSignal",
# use the longentry column as the sigcol
             arguments = list(sigcol = "longentry",
# set sigval to TRUE
                            sigval = TRUE,
# set orderqty to 1, all can only be used for exit rules
                            orderqty = 1,
# use a market type of order
                            ordertype = "market",
# take the long orderside
                            orderside = "long",
# do not replace other signals
                            replace = FALSE,
# buy at the next day's opening price
                            prefer = "Open"),
# this is an enter type rule, not an exit
             type = "enter")
```
  
  Running this RSI_3_4 strategy on the AES and the CHK over those 13 years, yields profit factors above 1, meaning that the strategy is profitable:

```{r, echo=FALSE, include = FALSE}
#### RUNNING YOUR STRATEGY ####
### Use applyStrategy() to apply your strategy. Save this to out
out <- applyStrategy(strategy = strategyRSI.st, portfolios = portfolioRSI.st)

### Update your portfolio (portfolio.st)
updatePortf(portfolioRSI.st)
daterange <- time(getPortfolio(portfolioRSI.st)$summary)[-1]

### Update your account (account.st)
updateAcct(accountRSI.st, daterange)
updateEndEq(accountRSI.st)

### Profit factor ########
#The profit factor is how many dollars you make for each dollar you lose. A profit factor above 1 means your strategy is profitable.
# Get the tradeStats for your portfolio
tstats <- tradeStats(Portfolios = portfolioRSI.st)

# Print the profit factor
#tstats$Profit.Factor
knitr::kable(tstats[, c("Symbol", "Profit.Factor")])
```
  
  Let's take a look at the system performance for the 2 stocks:

- Stock 1: AES  

```{r, echo=FALSE, include = FALSE}

#### USING CHART.POSN() ####
### use chart.Posn to view your system's performance on the stocks
#  This generates a crisp and informative visualization of the performance of your trading system over the course of the simulation.
# Chart trades against market data, position through time, and cumulative P\&L
# Produces a three-panel chart of time series charts that contains prices and transactions in the top panel, the resulting position in the second, and a cumulative profit-loss line chart in the third. 
# The percent positive statistic lets you know how many of your trades were winners. 

chart.Posn(Portfolio = portfolioRSI.st, Symbol = "stockOne")
### Adding indicators to a chart.Posn() chart ##########
# compute the SMA50
sma50 <- SMA(x = Cl(stockOne), n = 50)
# compute the SMA200
sma200 <- SMA(x = Cl(stockOne), n = 200)
# compute the DVO_2_126 with an navg of 2 and a percentlookback of 126
dvo <- DVO(HLC = HLC(stockOne), navg = 2, percentlookback = 126)
# recreate the chart.Posn of the strategy 
chart.Posn(Portfolio = portfolioRSI.st, Symbol = "stockOne")
# overlay the SMA50 on your plot as a blue line
add_TA(sma50, on = 1, col = "blue")
# overlay the SMA200 on your plot as a red line
add_TA(sma200, on = 1, col = "red")
# add the DVO_2_126 to the plot in a new window
```

```{r, echo=FALSE, message=FALSE}

add_TA(dvo)
```

  
- stock 2: CHK

```{r, echo=FALSE, message=FALSE, fig.keep = 'last'}

chart.Posn(Portfolio = portfolioRSI.st, Symbol = "stockTwo")
### Adding indicators to a chart.Posn() chart ##########
# compute the SMA50
sma50 <- SMA(x = Cl(stockOne), n = 50)
# compute the SMA200
sma200 <- SMA(x = Cl(stockOne), n = 200)
# compute the DVO_2_126 with an navg of 2 and a percentlookback of 126
dvo <- DVO(HLC = HLC(stockOne), navg = 2, percentlookback = 126)
# recreate the chart.Posn of the strategy 
chart.Posn(Portfolio = portfolioRSI.st, Symbol = "stockTwo")
# overlay the SMA50 on your plot as a blue line
add_TA(sma50, on = 1, col = "blue")
# overlay the SMA200 on your plot as a red line
add_TA(sma200, on = 1, col = "red")
# add the DVO_2_126 to the plot in a new window
add_TA(dvo)
```
  
  We can further analyze this RSI-3_4 strategy by getting the order book and retrieving the trade statistics:  


```{r, echo=FALSE, include = FALSE}
#### ANALYZE OUTPUT ####
# Order Book
getOrderBook(portfolioRSI.st)  

# Update Portfolio
updatePortf(portfolioRSI.st, Symbol = c("stockOne", "stockTwo"))

# Strategy output with trdeStats
tradeStats(portfolioRSI.st, Symbol = c("stockOne", "stockTwo"))
```

```{r, echo=FALSE, message = FALSE}
knitr::kable(t(tradeStats(portfolioRSI.st)))

#### COMPUTE SHARPE RATIO ####
### A Sharpe ratio is a metric that compares the average reward to the average risk taken. Generally, a Sharpe ratio above 1 is a marker of a strong strategy. 
# Get instrument returns
instretsRSI <- PortfReturns(portfolioRSI.st)

# Compute Sharpe ratio from returns
knitr::kable(SharpeRatio.annualized(instretsRSI, geometric = FALSE))
```
  
  
  For both instruments, the profit factor (absolute value ratio of gross profits over gross losses ) is above 1 . Therefore, this strategy is profitable.

Profit factor = Abs( gross profits / gross losses)  
  
  The sharpe ratio is a risk-adjusted measure of return.

Sharpe ratio = (Mean portfolio return - Risk-free rate ) / Standard deviation of portfolio return

  
  The annualized sharpe ratios are low, the highest being ~ 0.23 on stock AEP. Let's try to increase the annualized sharpe ratio by changing the oscillator of this strategy. 






## STRATEGY 2: DVO
  
  Instead of using the RSI_3_4 as the oscillator, let's use a custom DVO with navg = 2 and a percentlookback period of 126 that we call DVO_2_126 . The trend following indicators SMA50 and SMA200 stay the same
 , as well as the signals, rules and settings of the strategy.  


```{r, echo=FALSE, include = FALSE}

#### INITIALIZE STRATEGY ####
### Create initdate, from, and to charater strings
initdate <- "2000-01-01"
from <- "2003-01-01"
to <- "2016-12-31"

### Set the timezone to UTC
#OlsonNames()
Sys.setenv(TZ = "EST")

### Set the currency to USD 
currency("USD")

### Use the stock command to initialize and set currency to USD
stock(c("stockOne", "stockTwo"), currency = "USD")


### Define your trade size and initial equity
tradesize <- 100000
initeq <- 100000

### Define the names of your strategy, portfolio and account
strategyDVO.st <- "DVO strategy"
portfolioDVO.st <- "DVO strategy"
accountDVO.st <- "DVO strategy"

### Remove the existing strategy if it exists
rm.strat(strategyDVO.st)


### initialize the portfolio
initPortf(portfolioDVO.st, symbols = c("stockOne", "stockTwo"), initDate = initdate, currency = "USD")

### initialize the account
initAcct(accountDVO.st, portfolios = portfolioDVO.st, initDate = initdate, currency = "USD", initEq = initeq)

### initialize the orders
initOrders(portfolioDVO.st, initDate = initdate)

### store the strategy
strategy(strategyDVO.st, store = TRUE)



#### ADD SMA INDICATORS (TREND) ####
### 1. Add a 200-day simple moving average indicator to your strategy
add.indicator(strategy = strategyDVO.st,
# Add the SMA function
                  name = "SMA",
# Create a lookback period
                  arguments = list(x = quote(Cl(mktdata)), n = 200),
# Label your indicator SMA200
                  label = "SMA200")


### 2. Add a 50-day simple moving average indicator to your strategy
add.indicator(strategy = strategyDVO.st,
# Add the SMA function
                  name = "SMA",
# Create a lookback period
                  arguments = list(x = quote(Cl(mktdata)), n = 50),
# Label your indicator SMA50
                  label = "SMA50")


#### ADD OSCILLATOR DVO (David Varadi Oscillator, originated by David Varadi) ####
### Add the DVO indicator to your strategy
add.indicator(strategy = strategyDVO.st, name = "DVO",
              arguments = list(HLC = quote(HLC(mktdata)), navg = 2, percentlookback = 126),
              label = "DVO_2_126")


#### ADD SIGNALS ####
### 1. Using sigComparison ################
# add a sigComparison which specifies that SMA50 must be greater than SMA200, call it longfilter
add.signal(strategyDVO.st, name = "sigComparison",
# we are interested in the relationship between the SMA50 and the SMA200
               arguments = list(columns = c("SMA50", "SMA200"),
# particularly, we are interested when the SMA50 is greater than the SMA200
                                relationship = "gt"),
# label this signal longfilter
               label = "longfilter")

### 2. Using sigCrossover ###########
# add a sigCrossover which specifies that the SMA50 is less than the SMA200 and label it filterexit
add.signal(strategyDVO.st, name = "sigCrossover",
# we're interested in the relationship between the SMA50 and the SMA200
               arguments = list(columns = c("SMA50", "SMA200"),
# the relationship is that the SMA50 crosses under the SMA200
                                relationship = "lt"),
# label it filterexit
               label = "filterexit")


### 3.a Using sigThreshold - I ########
# SigThreshold() function takes the cross argument, which specifies whether it will function similarly to sigComparison (cross = FALSE) or sigCrossover (cross = TRUE), respectively.
# Implement a sigThreshold which specifies that DVO_2_126 must be less than 20, label it longthreshold
add.signal(strategyDVO.st, name = "sigThreshold",
# use the DVO_2_126 column
               arguments = list(column = "DVO_2_126",
# the threshold is 20
                                threshold = 20,
# we want the oscillator to be under this value
                                relationship = "lt",
# we're interested in every instance that the oscillator is less than 20
                                cross = FALSE),
# label it longthreshold
               label = "longthreshold")


### 3.b Using sigThreshold - II ###########
# add a sigThreshold signal to your strategy that specifies that DVO_2_126 must cross above 80 and label it thresholdexit
add.signal(strategyDVO.st, name = "sigThreshold",
# reference the column of DVO_2_126
               arguments = list(column = "DVO_2_126",
# set a threshold of 80
                                threshold = 80,
# the oscillator must be greater than 80
                                relationship = "gt",
# we are interested only in the cross
                                cross = TRUE),
# label it thresholdexit
               label = "thresholdexit")




### 4. Using sigFormula, Combining signals ############
# add a sigFormula signal to your code specifying that both longfilter and longthreshold must be TRUE, label it longentry
add.signal(strategyDVO.st, name = "sigFormula",
# specify that longfilter and longthreshold must be TRUE
               arguments = list(formula = "longfilter & longthreshold",
# specify that cross must be TRUE
                                cross = TRUE),
# label it longentry
               label = "longentry")

# Test
# Create your dataset: test
test_init <- applyIndicators(strategyDVO.st, mktdata = OHLC(stockOne))
test <- applySignals(strategy = strategyDVO.st, mktdata = test_init)
knitr::kable(head(test))

test_init2 <- applyIndicators(strategyDVO.st, mktdata = OHLC(stockTwo))
test2 <- applySignals(strategy = strategyDVO.st, mktdata = test_init2)
knitr::kable(head(test2))

### Subset your data between Sep. 1 and Sep. 5 of 2013
#test_subset <- test["2013-09-01/2013-09-05"]
#test_subset

#test_subset2 <- test2["2013-09-01/2013-09-05"]
#test_subset2




#### ADD RULES ####
### Using add.rule() to implement an exit rule #########
add.rule(strategyDVO.st, name = "ruleSignal",
         arguments = list(sigcol = "filterexit", sigval = TRUE, orderqty = "all",
                          ordertype = "market", orderside = "long",
                          replace = FALSE, prefer = "Open"),
         type = "exit")

#In quantstrat, the replace argument specifies whether or not to ignore all other signals on the same date when the strategy acts upon one signal.
### In this case, you will be working with a rule that sells when the DVO has crossed a certain threshold. In particular, you will be working with the thresholdexit rule now.
add.rule(strategyDVO.st, name = "ruleSignal",
         arguments = list(sigcol = "thresholdexit", sigval = TRUE, orderqty = "all",
                          ordertype = "market", orderside = "long",
                          replace = FALSE, prefer = "Open"),
         type = "exit")


### Using add.rule() to implement an entry rule #########
# create an entry rule of 1 share when all conditions line up to enter into a position
add.rule(strategyDVO.st, name = "ruleSignal",
# use the longentry column as the sigcol
             arguments = list(sigcol = "longentry",
# set sigval to TRUE
                            sigval = TRUE,
# set orderqty to 1, all can only be used for exit rules
                            orderqty = 1,
# use a market type of order
                            ordertype = "market",
# take the long orderside
                            orderside = "long",
# do not replace other signals
                            replace = FALSE,
# buy at the next day's opening price
                            prefer = "Open"),
# this is an enter type rule, not an exit
             type = "enter")


```
  




  Let's run the DVO strategy.

  
```{r, echo=FALSE, include = FALSE}

#### RUNNING YOUR STRATEGY ####
### Use applyStrategy() to apply your strategy. Save this to out
out <- applyStrategy(strategy = strategyDVO.st, portfolios = portfolioDVO.st)

### Update your portfolio (portfolio.st)
updatePortf(portfolioDVO.st)
daterange <- time(getPortfolio(portfolioDVO.st)$summary)[-1]

### Update your account (account.st)
updateAcct(accountDVO.st, daterange)
updateEndEq(accountDVO.st)

### what is the date of the last trade?
"2015-12-23"


### Profit factor ########
#The profit factor is how many dollars you make for each dollar you lose. A profit factor above 1 means your strategy is profitable.
# Get the tradeStats for your portfolio
tstats <- tradeStats(Portfolios = portfolioDVO.st)

# Print the profit factor
#tstats$Profit.Factor
tstats[, c("Symbol", "Profit.Factor")]
```

  
  Performance of the systems:  

```{r, echo=FALSE, message=FALSE, fig.keep = 'last'}
#### USING CHART.POSN() ####
### use chart.Posn to view your system's performance on stocks
#  This generates a crisp and informative visualization of the performance of your trading system over the course of the simulation.
# Chart trades against market data, position through time, and cumulative P\&L
# Produces a three-panel chart of time series charts that contains prices and transactions in the top panel, the resulting position in the second, and a cumulative profit-loss line chart in the third. 
# The percent positive statistic lets you know how many of your trades were winners. 

chart.Posn(Portfolio = portfolioDVO.st, Symbol = "stockOne")
### Adding indicators to a chart.Posn() chart ##########
# compute the SMA50
sma50 <- SMA(x = Cl(stockOne), n = 50)
# compute the SMA200
sma200 <- SMA(x = Cl(stockOne), n = 200)
# compute the DVO_2_126 with an navg of 2 and a percentlookback of 126
dvo <- DVO(HLC = HLC(stockOne), navg = 2, percentlookback = 126)
# recreate the chart.Posn of the strategy 
chart.Posn(Portfolio = portfolioDVO.st, Symbol = "stockOne")
# overlay the SMA50 on your plot as a blue line
add_TA(sma50, on = 1, col = "blue")
# overlay the SMA200 on your plot as a red line
add_TA(sma200, on = 1, col = "red")
# add the DVO_2_126 to the plot in a new window
add_TA(dvo)
```
  


```{r, echo=FALSE, message=FALSE, fig.keep = 'last'}
chart.Posn(Portfolio = portfolioDVO.st, Symbol = "stockTwo")
### Adding indicators to a chart.Posn() chart ##########
# compute the SMA50
sma50 <- SMA(x = Cl(stockOne), n = 50)
# compute the SMA200
sma200 <- SMA(x = Cl(stockOne), n = 200)
# compute the DVO_2_126 with an navg of 2 and a percentlookback of 126
dvo <- DVO(HLC = HLC(stockOne), navg = 2, percentlookback = 126)
# recreate the chart.Posn of the strategy 
chart.Posn(Portfolio = portfolioDVO.st, Symbol = "stockTwo")
# overlay the SMA50 on your plot as a blue line
add_TA(sma50, on = 1, col = "blue")
# overlay the SMA200 on your plot as a red line
add_TA(sma200, on = 1, col = "red")
# add the DVO_2_126 to the plot in a new window
add_TA(dvo)

```

  
  
  Analyzing this DVO strategy, we get the following trade metrics:  


```{r, echo=FALSE, message=FALSE, include = FALSE}
#### ANALYZE OUTPUT ####
# Order Book
getOrderBook(portfolioDVO.st) 
# Update Portfolio
updatePortf(portfolioDVO.st, Symbol = c("stockOne", "stockTwo"))

# Strategy output with trdeStats
tradeStats(portfolioDVO.st, Symbol = c("stockOne", "stockTwo"))
```

```{r, echo=FALSE, message=FALSE}
knitr::kable(t(tradeStats(portfolioDVO.st)))

#### COMPUTE SHARPE RATIO ####
### A Sharpe ratio is a metric that compares the average reward to the average risk taken. Generally, a Sharpe ratio above 1 is a marker of a strong strategy. 
# Get instrument returns
instretsDVO <- PortfReturns(portfolioDVO.st)

# Compute Sharpe ratio from returns
knitr::kable(SharpeRatio.annualized(instretsDVO, geometric = FALSE))
```
  

# CONCLUSION  
  
  On the same period of time, the same instruments/stocks and the strategy settings, the RSI strategy has a higher profit factor compared to the DVO strategy for each of the stocks respectively. 
  However, there are more transactions in the DVO strategy and its annualized sharpe ratios are much better than the ones of the RSI strategy. 
  Therefore, the absolute value of the gross profits over gross losses is higher in the RSI strategy for each respective stok, while the return per unit of risk is better in the DVO strategy.  
  
  Nonetheless, is that enough to select one strategy over the other? Would an entry rule with an order sizing functon instead of a single share considerably improve one strategy over 
  the other in terms of profit and risk-adjusted return?







