---
title: "DW-Exercise2"
author: "Diane K"
date: "September 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE}
##Load libraries
library(dplyr)
library(knitr)
```


##0. Load data to data frame
```{r message = FALSE}
titanic_dataset <- read.csv(file="titanic_original.csv"
                       , head = TRUE, sep=",")
str(titanic_dataset)

kable(head(titanic_dataset[,1:5]) , format = "markdown", caption = "Titanic dataset")
kable(head(titanic_dataset[,6:14]) , format = "markdown")
```


##1. Replace missing values in embarked with S
```{r message = FALSE}
if( is.factor(titanic_dataset$embarked)){
  levels(titanic_dataset$embarked)[1] <- "S"
}else{
  titanic_dataset$embarked[ is.na(titanic_dataset$embarked) |  
                              titanic_dataset$embarked == ""] <- "S" 
}
```


##2.a Replace missing age with mean age
```{r message = FALSE}
av_age <- mean( titanic_dataset$age, na.rm = TRUE)
titanic_dataset$age[ is.na(titanic_dataset$age) ] <- av_age
```

##2.b
- Could not replace with 0, because it's a wrong age for someone alive
- Could not replace with a negative value because it might impact calculations on the age column



##3. Replace missing values in LifeBoat column with "None"
```{r message = FALSE}
if( is.factor(titanic_dataset$boat)){
  levels(titanic_dataset$boat)[1] <- "None"
}else{
  titanic_dataset$boat[is.na(titanic_dataset$boat) ] <- "None"
}
```


##4. Add flag has_cabin_number
We should not replace missing cabin numbers because not all passengers purchased cabins

```{r message = FALSE}
titanic_dataset <- titanic_dataset %>% mutate(has_cabin_number = ifelse( 
  is.na(titanic_dataset$cabin) | (titanic_dataset$cabin) == "", 0, 1) ) 
```


##Output
```{r message = FALSE}
str(titanic_dataset)

kable(head(titanic_dataset[,1:5]) , format = "markdown", caption = "Titanic dataset cleaned")
kable(head(titanic_dataset[,6:12]) , format = "markdown") 
kable(head(titanic_dataset[,13:15]) , format = "markdown") 

write.csv(titanic_dataset, file = "titanic_clean.csv"
          , row.names = FALSE, append = FALSE)
```


