---
title: "Run Analysis on Smartphones Dataset"
author: "Diane K"
date: "September 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Load libraries
```{r  message = FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(pander)
```


## a. Merge training and test datasets into one dataset
```{r message = FALSE}
features <- read.table( 
  file = "./UCI HAR Dataset/features.txt") 
str(features)

kable( head(features),format = "markdown", caption = "Features" )
```

### Merge x training and test sets
```{r message = FALSE}
xTrain <- read.table( 
  file = "./UCI HAR Dataset/train/X_train.txt")
xTest <- read.table( 
  file = "./UCI HAR Dataset/test/X_test.txt")
x <- rbind(xTrain, xTest)

str(x)
```

### Name feature columns in x
```{r message = FALSE}
names(x) <- make.names(features$V2, unique = TRUE)
str(x)
pander(head(x), split.table = 80, style = 'rmarkdown')
```

### Read activities
```{r message = FALSE}
activityLabels <- read.table( 
  file = "./UCI HAR Dataset/activity_labels.txt") 
```

### Label activities
```{r message = FALSE}
names(activityLabels) <- c('ActivityLabel', 'ActivityName')
str(activityLabels)
kable( head(activityLabels),format = "markdown", caption = "Activities labels" )
```

### Merge subject train and test sets
```{r message = FALSE}
subjectTrain <- read.table( 
  file = "./UCI HAR Dataset/train/subject_train.txt") 
subjectTest <- read.table( 
  file = "./UCI HAR Dataset/test/subject_test.txt") 
subject <- rbind(subjectTrain, subjectTest)

names(subject) <- c('subject')
str(subject)
kable( head(subject),format = "markdown", caption = "Subjects" )
```

### Merge y training and test sets
```{r message = FALSE}
yTrain <- read.table( 
  file = "./UCI HAR Dataset/train/y_train.txt")
yTest <- read.table( 
  file = "./UCI HAR Dataset/test/y_test.txt")
y <- rbind(yTrain, yTest)

names(y) <- c('ActivityLabel')
str(y)
kable( head(y),format = "markdown", caption = "y" )
```

```{r message = FALSE, echo=FALSE}
# bodyAccXTrain <- read.table( 
#   file = "./UCI HAR Dataset/train/Inertial Signals/body_acc_x_train.txt")
# str(bodyAccXTrain)
# head(bodyAccXTrain)
```

### Merge all the datasets into 1
```{r message = FALSE}
smartphoneDataset <- cbind(subject, y) %>% left_join(activityLabels) %>% cbind(x)

str(smartphoneDataset)
pander(head(smartphoneDataset), split.table = 80, style = 'rmarkdown')
```



## b. Extract/remove columns containing mean and standard deviation for each variable
```{r message = FALSE}
smartphoneDatasetReduced <- smartphoneDataset %>% select( -matches("mean|std"))

str(smartphoneDatasetReduced)
pander(head(smartphoneDatasetReduced), split.table = 80, style = 'rmarkdown')
```


## c. Create tidy dataset with the average of each variable for each activity and each subject
```{r message = FALSE}
smartphoneDatasetAverages <- smartphoneDatasetReduced %>% 
  group_by(subject, ActivityName) %>%
  summarise_each(funs(mean), 
                 vars=-one_of(c('subject', 'ActivityLabel', 'ActivityName')) )

str(smartphoneDatasetAverages)
pander(head(smartphoneDatasetAverages), split.table = 80, style = 'rmarkdown')
```


## Output
```{r message = FALSE}
write.csv(smartphoneDataset, file = "smartphone_merged.csv"
          , row.names = FALSE, append = FALSE)
write.csv(smartphoneDatasetAverages, file = "smartphone_averages.csv"
          , row.names = FALSE, append = FALSE)
```
